---
title: "SeatCalculator - Highest Averages Methods v3.0"
author: ""
date: ""
output: 
  html_notebook: 
    fig_height: 9
    fig_width: 9
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



Explanations and code available here: <http://ismaelsb.github.io/ElectoralSpace> 

Fill in the boxes with the number of votes if you want to compute their allocated seats.

Fill in only the first boxes up to the number of parties specified, the rest will be ignored and not computed.

If not all, select a minimun of 3 parties to be shown from the specified before.

Select the 'step' value for the Electoral Method: (i.e. 0.5-Imperiali, 1-D'Hondt, 2-Sainte-Lague, 3-Danish).

```{r, echo = F, results='hide', message=FALSE, warning=FALSE}
#Uncommmet to install ggtern from an online CRAN repository:
#install.packages("ggtern")
#Load the ggtern library:
library(ggtern)
#library(stats)
#set.seed(156) #fix random generation

```

```{r, echo = F}

generateNodes <- function(seats){ 
  
  #define nodes
  nnodes = (seats+1)*(seats+2)/2;
  nodes  = matrix(0,nnodes,3);
  t=1;
  for (i in 0:seats){
    for (j in i:seats){
      nodes[t,] = c(i,j-i,seats-j);
      t=t+1;
    }
  }  
  
  index=as.matrix(1:nnodes)
  
  #node labels for seats
  label=cbind(as.character(nodes[,1]),matrix("-",nnodes),as.character(nodes[,2]),matrix("-",nnodes),as.character(nodes[,3]))
  label=do.call("paste0",as.data.frame(label))
  
  
  nodes <- cbind(index,as.data.frame(nodes),label)
  names(nodes) <- c("index",letters[24:26],"label")
  return(nodes);
  
}

```

```{r, echo = F}

generateColors <- function (colorRGB, seats) {
  
  nnodes = (seats+1)*(seats+2)/2;
  nodes <- as.matrix(generateNodes(seats)[,c("x","y","z")])
  #decimal codes for colors in nodes
  
  if (seats==0){values<-"#ffffff"} else {
  Code3 = floor(nodes*255.9/seats) #255.9 avoids seats -> 256 -> HEX #100 case
  #linear transformation to fixed extreme colors
  CodeDecRGB = floor(Code3%*%matrix(colorRGB,3,3, byrow=T)/256)
  #hex codes for colors in nodes
  CodeRGB=cbind(matrix("#",nnodes),format(as.hexmode(CodeDecRGB[,1]),width=2),format(as.hexmode(CodeDecRGB[,2]),width=2),format(as.hexmode(CodeDecRGB[,3]),width=2))
  CodeRGB=do.call("paste0",as.data.frame(CodeRGB))
  values=CodeRGB
  }
  
  return(values);
  
}

```

```{r, echo = F}

alloc <- function(parties, votes, seats, step, threshold=0){
  
  #function for the seat allocation and its ordering
  #computes allocation for several values of seats and divisor step methods
  #with the same quotiens table


  
  votes=votes*(votes>=(threshold*sum(votes)))
  
  if (sum(votes>0)==0) {
    
       votes <- votes*0+1
       #to avoid an error in the share of seats
    
  }
  
  Mst <- max(step)
  Mse <- max(seats,1) #1 to avoid 0
  nquotients=(1+Mst*(Mse-1))
  
  #table with all the quotients needed
  quotienstable <- data.frame( 
    parties    = rep(parties, each = nquotients), 
    quotients  = as.vector(sapply(votes, function(x) 
      x/seq(from=1, to=nquotients) )),
    votesrep   = rep(votes, each = nquotients), 
    #in a tie of quotiens, the party with more votes first
    tiebreak = rep(runif(length(parties)), each = nquotients) #
    #in a tie of votes, first seats allocated at random quotientsthen the nexts seats in the same order
  ) 
  
  SeatsList=list()
  filteredtable =list()
  
  
  for (j in 1:length(step)) {
    
    
    #select from the table of all quotiens the ones whose divisors
    #belong to the current sequence, given 'seats' and 'step'
    select <- rep(seq(from=1, to=1+step[j]*(Mse-1), by=step[j]),length(parties))+
      rep(seq(from=0,to=nquotients*length(parties)-1,by=nquotients),each=Mse)
    
    filteredtable [[j]] <- quotienstable$parties[select][order(-quotienstable$quotients[select], -quotienstable$votesrep[select],quotienstable$tiebreak[select])] #,runif(length(parties))
    
    SeatsList[[j]]<-list()
    
    #a vector of values for seats if you want to compute the partial sums
    for (i in 1:length(seats)){
      
      SeatsList[[j]][[i]]= table(filteredtable[[j]][1:seats[i]])
      
      if (seats[i]==0){
    
        SeatsList[[j]][[i]]=0*SeatsList[[j]][[i]]
        
    
      }

      
    }
    
    names(SeatsList[[j]])[1:length(seats)]<-do.call("paste0",as.data.frame(cbind("divisor step ",step[j]," for ", seats," seats")))
    
    
    SeatsList[[j]][[length(seats)+1]] = as.matrix(filteredtable[[j]][1:Mse]) #ordering
    
    names(SeatsList[[j]])[length(seats)+1]<-paste0("ordering for divisor step ", step[j]," for ", Mse," seats")
    
    
  }
  
  
  return(SeatsList);
  
}

```

```{r, echo = F}

AllocatedNode <- function (y, nodes, nnodes) {
  
  Node = which.min(apply(abs(nodes - rep(1,nnodes) %*% t(y)),1,max))
  
  return(Node);
  
}

```

```{r, echo = F}

generateHexagonData <- function(seats, step=1) {
  
  divisor <- seq(from=1, to=1+step*(seats+2-1), by=step)
  divisor <- c(0,divisor)
  
  NodesData <- generateNodes(seats)
  
  hexagon <- matrix(0,nrow=(seats+1)*(seats+2)/2, ncol=6*3)
  
  for (i in 1:((seats+1)*(seats+2)/2)){
    
    n <- NodesData[i,2:4] #node
    hexagon[i,] <- rep(c(n[[1]],n[[2]],n[[3]]),6)+c(1,0,0, 1,0,1, 0,0,1, 0,1,1, 0,1,0, 1,1,0)
    hexagon[i,] <- divisor[hexagon[i,]+1]
    
  }
  
  dhex <- as.data.frame(cbind(rep(1:((seats+1)*(seats+2)/2),each=6), matrix(t(hexagon),nrow=6*(seats+1)*(seats+2)/2, ncol=3, byrow=T),
                              rep(NodesData$x,each=6),rep(NodesData$y,each=6),rep(NodesData$z,each=6)))
  
  names(dhex) <- c('hex','x','y','z','nx','ny','nz')
  
  return(dhex)
  
}

```

```{r, echo = F}

generatelabels <- function(seats, step=1){
  
  #labels for the center of the regions
  
  nnodes     <- (seats+1)*(seats+2)/2
  nodes      <- generateNodes(seats)
  dhex       <- generateHexagonData(seats, step=step)
  labelsData <- as.data.frame(matrix(0,nnodes, 5))
  
  for (index in 1:nnodes) {
    
    labelsData[index,1]   <- as.integer(index)
    labelsData[index,2:4] <- as.numeric(t(colSums(dhex[dhex$hex==index,2:4])))
    labelsData[index,5]   <- toString(nodes[index,5])
    
  }
  
  names(labelsData) <- c("index","x","y","z","label")
  
  return(labelsData)
  
}

```

```{r, echo = F}

plotallocation <- function (seats, step=1, colorRGB0=c(242,74,87, 124,218,198, 2,114,195), alpha=0.75, labelsize=4, dotsize=1, dotcolor='khaki2') {
  
  NodesData <- generateNodes(seats)
  dhex <- generateHexagonData(seats, step)
  labelsData <- generatelabels(seats, step)
  
  polygons <- ggtern(data=dhex, aes(x, y, z)) + theme_nomask()
  
  for (i in 1:((seats+1)*(seats+2)/2)){
    
    polygons <- polygons + geom_polygon(aes(fill=hex, group=hex), data=dhex[dhex==i,] , fill=generateColors(colorRGB0,seats)[i], colour='grey40', size=0, alpha=alpha) #no lines
    
  }
  
  polygons <- polygons + geom_point(data=NodesData,aes(x,y),color=dotcolor, size=dotsize)+
    #geom_text(data=NodesData, aes(label=label), color="grey30", hjust=0.5, vjust=-0.6, size=4)+ 
    labs(x="X",y="Y",z="Z",title="")
  
  polygons <- polygons + #geom_point(data=labelsData,aes(x,y,z),color="khaki2")+
    geom_text(data=labelsData, aes(label=label), color="grey30", hjust=0.5, vjust=0.5, size=labelsize)+ 
    labs(x="X",y="Y",z="Z",title="")
  
  return(polygons)
  
}

```

```{r, echo = F}

plotSeatCalculator<-function(seats,step,votes1,votes2,votes3,color1,color2,color3,Opacity=1,labelsize=4,dotsize=2,dotcolor='khaki2',threshold=0){
  
  
  
  colorRGB0 <- matrix(rbind(col2rgb(color1),
                            col2rgb(color2),
                            col2rgb(color3)),1,9)
  
  
  TQp <- plotallocation(seats, step=step, colorRGB0=colorRGB0, alpha=Opacity, labelsize=labelsize, dotsize=dotsize, dotcolor=dotcolor)
  
  
  NodesData <- generateNodes(seats)
  
  
  votes <- as.data.frame(cbind(0,0,0))
  votesinput <- as.data.frame(cbind(votes1,votes2,votes3))
  names(votes) <- c("x","y","z")
  names(votesinput) <- c("x","y","z")
  
  
  
  if (max(votesinput)>0 & is.na(min(votesinput))==F) {
    
    if (seats==0) {
    
      votes <- votesinput
    
      TQp <- TQp + geom_point(data=votes ,aes(x,y),color="grey20", size=dotsize+2, alpha=.8)
    
    } else {
    
    votes <- votesinput
    
    allocated <- alloc(letters[24:26], unlist(votes), seats, step, threshold)
    
    nodes  <- as.matrix(NodesData[,c("x","y","z")])
    nnodes <- (seats+1)*(seats+2)/2
    
    index <- AllocatedNode(allocated[[1]][[1]], nodes, nnodes)
    
    dhex <- generateHexagonData(seats, step=step)
    
    labelcoords <- as.data.frame(t(colSums(dhex[dhex$hex==index,2:4])))
    labelcoords <- cbind(labelcoords,NodesData[index,]$label)
    names(labelcoords)[4] <- "label"
    
    
    TQp <- TQp + geom_polygon(aes(fill=hex, group=hex), data=dhex[dhex==index,] , fill=generateColors(colorRGB0,seats)[index], colour='grey20', size=1, alpha=0.85) +
      geom_point(data=votes ,aes(x,y),color="grey20", size=dotsize+2, alpha=.8)+
      geom_text(data=labelcoords,aes(label=label), color="grey15", hjust=0.5, vjust=0.5, size=labelsize+2)+
      geom_point(data=NodesData[index,],aes(x,y),color=dotcolor,size=dotsize)
    } 
  
  }
  
  
  
  return(TQp)

  
}

```

```{r echo=FALSE, warning=FALSE}
inputPanel(
  
  #radioButtons("step", label = "Number of steps:", 1:3, selected=1, inline=T),
  
  #radioButtons("electmethod", label = "Electoral method:", c("D'Hondt","Sainte-Lague", "Danish"), selected="D'Hondt", inline=F),
  
  numericInput("seats", label="Number of seats:", value=4, min = 0, step=1, width='75%'),
  
  numericInput("st", label="Step value:", min = 0.25, max = 5, value = 1, step = 0.25, width='75%'),
  
  #sliderInput("st", label = "Step value:", min = 0.25, max = 5, value = 1, step = 0.25),
    
  #sliderInput("seats", label = "Number of seats:",
  #            min = 1, max = 131, value = 4, step = 1),
  
  sliderInput("n", label = "Number of parties:", min = 3, max = 8, value = 3, step = 1, width='75%')
  
  #sliderInput("threshold", label = "Threshold:",
  #            min = 0, max = .30, value = 0, step = .01)
  
  #actionButton("button", "Compute!")
  
  #checkboxGroupInput("selectparties", label="Show parties:", choiceNames = LETTERS[1:8], choiceValues = 1:8 , selected = 1:8, inline = T)
  
  
)

inputPanel(
  
  numericInput("votes1", label="Votes for Party A", value=0, min = 0),
  numericInput("votes2", label="Votes for Party B", value=0, min = 0),
  numericInput("votes3", label="Votes for Party C", value=0, min = 0),
  numericInput("votes4", label="Votes for Party D", value=0, min = 0),
  numericInput("votes5", label="Votes for Party E", value=0, min = 0),
  numericInput("votes6", label="Votes for Party F", value=0, min = 0),
  numericInput("votes7", label="Votes for Party G", value=0, min = 0),
  numericInput("votes8", label="Votes for Party H", value=0, min = 0)
  
)




renderPlot({

  
  #observeEvent(input$button, {
  #eventReactive(input$button,{  
    
  step <- input$st
  
  #if (input$electmethod == "D'Hondt"){
    
  #  step <- 1
    
  #} else if (input$electmethod == "Sainte-Lague"){
    
  #  step <- 2
    
  #} else if (input$electmethod == "Danish"){
    
  #  step <- 3
    
  #}
  
  
  votesvector0 <- as.vector(c(input$votes1,input$votes2,input$votes3,input$votes4,input$votes5,input$votes6,input$votes7,input$votes8))
  
  #votesvector <- votesvector0[votesvector0>0]
  
  #n <- sum(votesvector>0)
  
  
  
  n <- input$n
  votesvector <- votesvector0[1:n]
  
  colorinput0 <- c(input$color1,input$color2,input$color3,input$color4,input$color5,input$color6,input$color7,input$color8)
  colorinput <- colorinput0[1:n]
  
  
  partiessubsets <- t(combn(n,3))
  ncombparties <- dim(partiessubsets)[1]
  alphabet <- LETTERS[1:n]
  
  

  allocTotal <- alloc(alphabet, votesvector, input$seats, step, threshold=0) 


  #select parties
  #selectparties <- (1:n)[input$selectparties[input$selectparties<=n]] #selected from the first n
  #allocSelect <- allocTotal[[1]][[1]][selectparties]
  #seatsSelect <- sum(allocSelect)
  
  #s <- length(selectparties)
  
  
  
  if (n>=3) {
  
  partiessubsets <- t(combn(n,3))
  ncombparties <- dim(partiessubsets)[1]
  #
  
  
     
  TQp <- list()
  
  for (i in 1:ncombparties){
    
    allocSubset <- allocTotal[[1]][[1]][partiessubsets[i,]]
    seatsSubset <- sum(allocSubset)
    
  
    TQp[[i]] <- plotSeatCalculator(seatsSubset,step,votesvector[partiessubsets[i,1]],votesvector[partiessubsets[i,2]],votesvector[partiessubsets[i,3]],colorinput[partiessubsets[i,1]],colorinput[partiessubsets[i,2]],colorinput[partiessubsets[i,3]],input$Opacity,input$labelsize,input$dotsize,input$dotcolor,threshold=0)+
      xlab(alphabet[partiessubsets[i,1]])+
      ylab(alphabet[partiessubsets[i,2]])+
      zlab(alphabet[partiessubsets[i,3]])
    
    
    
  }

    #c <-round(sqrt(n*(n-1)*(n-2)/12)) #num of columns: x*2x rectangle ~ (n comb 3) number of plots
    #c <- 2

    plist=TQp
    multiTQp <- grid.arrange(grobs = plist)#, ncol=c)
    multiTQp 
    
  }#end  if s >=3 
    
  #})
    
  #}, autoDestroy=T)
  
  
  
    
  
    

    
  }, width=2800, height=2800, res=180)#, width=2800, height=5600, res=180)#, width=3600, height=2800, res=288
  


inputPanel(
  
  sliderInput("Opacity", label = "Opacity:",
              min = 0, max = 1, value = .7, step = .1),
  
  sliderInput("labelsize", label = "Label size:",
              min = 0, max = 12, value = 4, step = 1),
  
  sliderInput("dotsize", label = "Dot size:",
              min = 0, max = 5, value = 2, step = 1),
  
  selectInput("dotcolor", label = "Dot color:",
              choices = c('red','green','blue','yellow','magenta','cyan','purple','brown','orange','antiquewhite','aquamarine3','azure1','brown3','brown4','burlywood4','cadetblue','coral3','cornsilk','cornsilk2','cyan3','darkcyan','darkolivegreen4','darkorange2','darkred','darkslategray','darkseagreen','darksalmon','darkslategray3','deeppink4','dodgerblue3','darkorchid2','dodgerblue4','firebrick3','floralwhite','honeydew3','hotpink4','khaki2','khaki3','khaki4','lavenderblush3','lightblue3','lightcyan2','lightpink2','lightsalmon2','lightsteelblue3','lightyellow1','indianred3','indianred4','ivory','ivory3','midnightblue','mintcream','mistyrose2','navajowhite1','orangered4','paleturquoise2','palegreen3','rosybrown2','royalblue3','salmon','seagreen','seashell1','sienna2','turquoise3','wheat2','black','gray10','gray30','gray50','gray70','gray90','gray','white'), selected = 'khaki2')
  
)


inputPanel(
  
  selectInput("color1", label = "Color A:",
              choices = c('red','green','blue','yellow','magenta','cyan','purple','brown','orange','antiquewhite','aquamarine3','azure1','brown3','brown4','burlywood4','cadetblue','coral3','cornsilk','cornsilk2','cyan3','darkcyan','darkolivegreen4','darkorange2','darkred','darkslategray','darkseagreen','darksalmon','darkslategray3','deeppink4','dodgerblue3','darkorchid2','dodgerblue4','firebrick3','floralwhite','honeydew3','hotpink4','khaki3','khaki4','lavenderblush3','lightblue3','lightcyan2','lightpink2','lightsalmon2','lightsteelblue3','lightyellow1','indianred3','indianred4','ivory','ivory3','midnightblue','mintcream','mistyrose2','navajowhite1','orangered4','paleturquoise2','palegreen3','rosybrown2','royalblue3','salmon','seagreen','seashell1','sienna2','turquoise3','wheat2','black','gray10','gray30','gray50','gray70','gray90','gray','white'), selected = 'yellow'),
  
  selectInput("color2", label = "Color B:",
              choices = c('red','green','blue','yellow','magenta','cyan','purple','brown','orange','antiquewhite','aquamarine3','azure1','brown3','brown4','burlywood4','cadetblue','coral3','cornsilk','cornsilk2','cyan3','darkcyan','darkolivegreen4','darkorange2','darkred','darkslategray','darkseagreen','darksalmon','darkslategray3','deeppink4','dodgerblue3','darkorchid2','dodgerblue4','firebrick3','floralwhite','honeydew3','hotpink4','khaki3','khaki4','lavenderblush3','lightblue3','lightcyan2','lightpink2','lightsalmon2','lightsteelblue3','lightyellow1','indianred3','indianred4','ivory','ivory3','midnightblue','mintcream','mistyrose2','navajowhite1','orangered4','paleturquoise2','palegreen3','rosybrown2','royalblue3','salmon','seagreen','seashell1','sienna2','turquoise3','wheat2','black','gray10','gray30','gray50','gray70','gray90','gray','white'), selected = 'magenta'),
  
  selectInput("color3", label = "Color C:",
              choices = c('red','green','blue','yellow','magenta','cyan','purple','brown','orange','antiquewhite','aquamarine3','azure1','brown3','brown4','burlywood4','cadetblue','coral3','cornsilk','cornsilk2','cyan3','darkcyan','darkolivegreen4','darkorange2','darkred','darkslategray','darkseagreen','darksalmon','darkslategray3','deeppink4','dodgerblue3','darkorchid2','dodgerblue4','firebrick3','floralwhite','honeydew3','hotpink4','khaki3','khaki4','lavenderblush3','lightblue3','lightcyan2','lightpink2','lightsalmon2','lightsteelblue3','lightyellow1','indianred3','indianred4','ivory','ivory3','midnightblue','mintcream','mistyrose2','navajowhite1','orangered4','paleturquoise2','palegreen3','rosybrown2','royalblue3','salmon','seagreen','seashell1','sienna2','turquoise3','wheat2','black','gray10','gray30','gray50','gray70','gray90','gray','white'), selected = 'cyan'),
  
  selectInput("color4", label = "Color D:",
              choices = c('red','green','blue','yellow','magenta','cyan','purple','brown','orange','antiquewhite','aquamarine3','azure1','brown3','brown4','burlywood4','cadetblue','coral3','cornsilk','cornsilk2','cyan3','darkcyan','darkolivegreen4','darkorange2','darkred','darkslategray','darkseagreen','darksalmon','darkslategray3','deeppink4','dodgerblue3','darkorchid2','dodgerblue4','firebrick3','floralwhite','honeydew3','hotpink4','khaki3','khaki4','lavenderblush3','lightblue3','lightcyan2','lightpink2','lightsalmon2','lightsteelblue3','lightyellow1','indianred3','indianred4','ivory','ivory3','midnightblue','mintcream','mistyrose2','navajowhite1','orangered4','paleturquoise2','palegreen3','rosybrown2','royalblue3','salmon','seagreen','seashell1','sienna2','turquoise3','wheat2','black','gray10','gray30','gray50','gray70','gray90','gray','white'), selected = 'red'),
    
  selectInput("color5", label = "Color E:",
              choices = c('red','green','blue','yellow','magenta','cyan','purple','brown','orange','antiquewhite','aquamarine3','azure1','brown3','brown4','burlywood4','cadetblue','coral3','cornsilk','cornsilk2','cyan3','darkcyan','darkolivegreen4','darkorange2','darkred','darkslategray','darkseagreen','darksalmon','darkslategray3','deeppink4','dodgerblue3','darkorchid2','dodgerblue4','firebrick3','floralwhite','honeydew3','hotpink4','khaki3','khaki4','lavenderblush3','lightblue3','lightcyan2','lightpink2','lightsalmon2','lightsteelblue3','lightyellow1','indianred3','indianred4','ivory','ivory3','midnightblue','mintcream','mistyrose2','navajowhite1','orangered4','paleturquoise2','palegreen3','rosybrown2','royalblue3','salmon','seagreen','seashell1','sienna2','turquoise3','wheat2','black','gray10','gray30','gray50','gray70','gray90','gray','white'), selected = 'green'),
    
  selectInput("color6", label = "Color F:",
              choices = c('red','green','blue','yellow','magenta','cyan','purple','brown','orange','antiquewhite','aquamarine3','azure1','brown3','brown4','burlywood4','cadetblue','coral3','cornsilk','cornsilk2','cyan3','darkcyan','darkolivegreen4','darkorange2','darkred','darkslategray','darkseagreen','darksalmon','darkslategray3','deeppink4','dodgerblue3','darkorchid2','dodgerblue4','firebrick3','floralwhite','honeydew3','hotpink4','khaki3','khaki4','lavenderblush3','lightblue3','lightcyan2','lightpink2','lightsalmon2','lightsteelblue3','lightyellow1','indianred3','indianred4','ivory','ivory3','midnightblue','mintcream','mistyrose2','navajowhite1','orangered4','paleturquoise2','palegreen3','rosybrown2','royalblue3','salmon','seagreen','seashell1','sienna2','turquoise3','wheat2','black','gray10','gray30','gray50','gray70','gray90','gray','white'), selected = 'blue'),

    selectInput("color7", label = "Color G:",
              choices = c('red','green','blue','yellow','magenta','cyan','purple','brown','orange','antiquewhite','aquamarine3','azure1','brown3','brown4','burlywood4','cadetblue','coral3','cornsilk','cornsilk2','cyan3','darkcyan','darkolivegreen4','darkorange2','darkred','darkslategray','darkseagreen','darksalmon','darkslategray3','deeppink4','dodgerblue3','darkorchid2','dodgerblue4','firebrick3','floralwhite','honeydew3','hotpink4','khaki3','khaki4','lavenderblush3','lightblue3','lightcyan2','lightpink2','lightsalmon2','lightsteelblue3','lightyellow1','indianred3','indianred4','ivory','ivory3','midnightblue','mintcream','mistyrose2','navajowhite1','orangered4','paleturquoise2','palegreen3','rosybrown2','royalblue3','salmon','seagreen','seashell1','sienna2','turquoise3','wheat2','black','gray10','gray30','gray50','gray70','gray90','gray','white'), selected = 'white'),

    selectInput("color8", label = "Color H:",
              choices = c('red','green','blue','yellow','magenta','cyan','purple','brown','orange','antiquewhite','aquamarine3','azure1','brown3','brown4','burlywood4','cadetblue','coral3','cornsilk','cornsilk2','cyan3','darkcyan','darkolivegreen4','darkorange2','darkred','darkslategray','darkseagreen','darksalmon','darkslategray3','deeppink4','dodgerblue3','darkorchid2','dodgerblue4','firebrick3','floralwhite','honeydew3','hotpink4','khaki3','khaki4','lavenderblush3','lightblue3','lightcyan2','lightpink2','lightsalmon2','lightsteelblue3','lightyellow1','indianred3','indianred4','ivory','ivory3','midnightblue','mintcream','mistyrose2','navajowhite1','orangered4','paleturquoise2','palegreen3','rosybrown2','royalblue3','salmon','seagreen','seashell1','sienna2','turquoise3','wheat2','black','gray10','gray30','gray50','gray70','gray90','gray','white'), selected = 'gray')
    
)







 




